---
title: "McDonalds Nutritional Analysis"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

This Report shows the 

Download the[dataset here](https://www.kaggle.com/datasets/mcdonalds/nutrition-facts/data)


*Load the Packages for data manipulation and visualization*

```{r}
library(dplyr)
library(tidyverse)
library(GGally)
```
**Read in the data**


```{r}

menu <- read.csv("menu.csv")


```


**How many Menu's do each Category have?**


*This gives a quick glance at the data before we begin exploration*
```{r}
cat <- menu %>% 
  select(Category) %>% 
  count(Category,name = "Total")  %>% 
  distinct()

cat_plot <- cat %>% 
  ggplot(aes(Category,
             Total,
             fill = Category)) + 
  geom_col() + 
  theme_classic() + 
  labs(title = "Total Categories") + 
  theme(legend.position = "none") + 
  geom_text(aes(label = Total), 
            position = position_stack(vjust = 0.5),
            color ="black", size = 3) + coord_flip()


cat_plot

#Coffee & Tea offers a lot a variation on the menu

```


**Average Protein Content in the Menu Category**
```{r}
Pro_avg <- menu %>% select(Category,Item,Protein) %>% group_by(Category) %>%
  aggregate(Protein~Category, FUN = mean)

Pro_avg


#Chicken & Fish, Beef & pork contain higher protein content in every menu item
```


**What Menu item contains the highest amount of Protein content?**
```{r}
max_pro <- menu %>% 
  select(Item,Protein) %>%  
  group_by(Item) %>% 
  arrange(desc(Protein)) %>% 
  head(15) %>% 
  ggplot(aes(x= reorder(Item, 
                        Protein),
             fill = Item)) + 
  geom_bar(stat = "identity",
           aes(y= Protein )) +
  coord_flip() +
  theme_bw()+ 
  geom_text(aes(label = Protein, 
                y= Protein),
            position = position_stack(vjust = 0.5),
            color = "Black",
            size = 3)+
  theme(legend.position="none") +
  ggtitle("TOP 15 McDonald's Menu Item with the Highest Protein Content") +theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())

max_pro

#Chicken McNuggents(40 pieces) tops the chart for meals with the highest protein content.
```


**What Menu item contains the least amount of Protein content?**
```{r}
min_pro <- menu %>% 
  select(Item,Protein) %>%  
  group_by(Item, Protein) %>% 
  filter(Protein >0 )%>% 
  arrange(Protein) %>% 
  head(15)  %>% 
  ggplot(aes(x= reorder(Item, 
                        Protein),
             fill = Item)) + 
  geom_bar(stat = "identity",
           aes(y= Protein )) +
  coord_flip() +
  theme_bw()+ 
  geom_text(aes(label = Protein, 
                y= Protein),
            position = position_stack(vjust = 0.5),
            color = "Black",
            size = 3)+
  ggtitle("TOP 10 McDonald's Menu Item with the Least Protein")+
  theme(legend.position="none")

min_pro

#The protein levels in this category are very low.
```


**How do the Menu category's rank in order of Calories?**
```{r}
cal <- menu %>% 
  select(Category,
         Item,
         Calories) %>% 
  group_by(Category)  



cal_plot <- cal %>% 
  ggplot(aes(Calories,
             Category,
             fill = Category)) + 
  geom_boxplot() + 
  theme_classic() + 
  scale_fill_brewer(palette = "Set1") + ggtitle("Calorie Content per Menu Category")

cal_plot

#Smoothies & Shakes are going to give you the energy you need!!!
```


**Top 10 Menu Food items with High calories**
```{r}
max_cal <- menu %>% 
  select(Category,
         Item,
         Calories) %>% 
  group_by(Item) %>% 
  arrange(desc(Calories))%>% 
  head(10)%>% 
  ggplot(aes(Calories,
             y = reorder(Item,
                         Calories)
             ,fill =Item)) + 
  geom_col() +
  theme_classic()+
  theme(legend.position = "none") + 
  geom_text(aes(label = Calories), 
            position = position_stack(vjust = 0.5),
            color="black", size =3) + ggtitle("TOP 10 McDonald's Menu Item with the Highest Calorie Content")

max_cal

#Proteinous and Energy packed? Power to the McNuggets(40 pieces) !!!
```


**Top 10 Menu Food items with Low calories**
```{r}
min_cal <- menu %>%
  select(Category,
         Item,
         Calories)%>% 
  filter(Calories>0) %>% 
  group_by(Item) %>% 
  arrange(Calories)%>% 
  head(10) %>% 
  ggplot(aes(Calories,
             y = reorder(Item,
                         Calories),
             fill =Item)) + 
  geom_col() + 
  theme_classic()+ 
  theme(legend.position = "none") + 
  geom_text(aes(label = Calories), 
            position = position_stack(vjust = 0.5),
            color="black", size =3) +
  scale_fill_brewer(palette = "Set3") + ggtitle("TOP 10 McDonald's Menu Item with Low Calorie Content")

min_cal

#That's not a lot energy  but it's energy nonetheless.

```


**Menu Category with the Highest Carbohydrates Content**
```{r}
carb <- menu %>% 
  select(Category,
         Item,
         Carbohydrates)%>% 
  group_by(Category) 

carb_plot<- carb %>%
  ggplot(aes(Category,Carbohydrates, 
             fill = Category)) + 
  geom_boxplot() +
  scale_fill_brewer(palette = "Set1") + coord_flip() + theme_classic() + ggtitle("Menu Category with the Highest Carbohydrates Content")

carb_plot

# Carbohydrates supply your body with glucose, which your cells use as fuel. Certain complex carbohydrates, like fiber, also feed your gut microbiome.
```


**Top 10 Menu Food items with High Sugar Content**
```{r}
sugar <- menu %>%
  select(Category,Item,Sugars) %>% 
  group_by(Category)

max_sugar <-sugar %>% 
  arrange(desc(Sugars)) %>% 
  head(10) %>% 
  ggplot(aes(Sugars,
             y=reorder(Item,
                       Sugars),
             fill = Sugars)) +
  geom_col() + 
  theme_classic()+
  theme(legend.position = "none") +
  geom_text(aes(label = Sugars),
            position = position_stack(vjust = 0.5)) + ggtitle("Top 10 Menu items with the Highest Sugar Content")


max_sugar

#Smoothies & Shakes come out the cream of crop in this category.
```


**Top 10 Menu Food items with Low Sugar Content**
```{r}
min_sugar <- sugar %>% 
  filter(Sugars>0) %>% 
  arrange(Sugars) %>% 
  head(10) %>% 
  ggplot(aes(Sugars, 
             y=reorder(Item,
                       Sugars),
             fill = Sugars)) +
  geom_col() + 
  theme_classic()+
  theme(legend.position = "none") + 
  geom_text(aes(label = Sugars),
            position = position_stack(vjust = 0.5)) + ggtitle("Top 10 Menu items with the Lowest sugar Content")


min_sugar

#Apparently chicken isn't sweet but tasty. Oh well better luck next time Mr. McNuggets!
```



**Menu Categories with Carbohydrates that contain a higher Fibre to Sugar Ratio** 
```{r}
healthy_carbs <-  menu %>% 
  select(Carbohydrates,
         Category,
         Sugars,
         Dietary.Fiber)  %>%  
  filter(Dietary.Fiber > Sugars)

healthy_carbs_plot <- healthy_carbs %>% 
  ggplot(aes(Carbohydrates,
             y = reorder(Category,Carbohydrates),
             fill = Category)) +
  geom_boxplot() + ggtitle("Menu Categories Rich in Carbohydrates and Fibre") + theme_classic()



healthy_carbs_plot

#As a general rule, the more fiber and the less sugar there is in the carbohydrate, the better it is for your body.

```



**Menu Categories grouped by sodium content**
```{r}
sodium  <- menu %>%
  select(Category,
         Item,Sodium,
         Sodium....Daily.Value.) %>% 
  group_by(Category)


sodium_plot <- sodium %>% 
  ggplot(aes(Sodium,
             Category,
             fill = Category)) +
  geom_boxplot() + theme_classic() + ggtitle("Menu Categories grouped by sodium content")


sodium_plot

# Our bodies need sodium; however, there is so much in most processed foods, and too much sodium can have several negative effects. Canned foods are notorious for having lots of sodium, as are bottled sauces, salty snacks, and processed foods. Most nutritionists recommend watching your sodium intake carefully because too much sodium in your diet is dangerous. It can lead to high blood pressure and many related health conditions.
```


**Menu Categories grouped by Trans Fat**
```{r}
trans_fat <- menu %>%
  select(Category,
         Trans.Fat) %>%
  group_by(Category)

trans_plot <- trans_fat %>% 
  ggplot(aes(Trans.Fat,
             Category,
             fill = Category)) + 
  geom_boxplot() + theme_classic() +  ggtitle("Menu Categories grouped by trans_fat content")

trans_plot

# In the United States, the main dietary source of trans fats is partially hydrogenated vegetable oils, previously used in many commercially prepared foods. Consuming trans fats may adversely affect cholesterol levels in the body and may contribute to the risk of atherosclerosis. Because of this, the US Food and Drug Administration (FDA) has banned the use of trans fats in prepared foods.
```


**By how much do Protein,Carbohydrates and Fat affect the Calories?**
```{r}
calories <- menu %>%
  select(Calories,
         Carbohydrates, 
         Protein,
         Total.Fat) %>% 
  group_by(Calories)

calories_plot <- ggcorr(calories,
       method =  c("everything", "pearson"),
       label = TRUE)  + theme_classic() +  ggtitle("How Protein,Carbohydrates and Fat contribute to Calories?")

calories_plot

#Fats are the slowest source of energy but the most energy-efficient form of food.Because fats are such an efficient form of energy, the body stores any excess energy as fat
```

